### 2014-10-23 paper/include/background.tex

\todo[inline]{change the next paragraph to introduce retention time.}
The memory controller (MC) normally issues refresh operations periodically at an interval of \textit{64~ms} for each row, an interval that has been constant for the latest DRAM specifications [source] and called \textit{maximum refresh period}. This interval is conservatively halved if the DRAM exceeds the specified working temperature of $85^{\circ}C$.

>>>>

The memory controller (MC) normally issues refresh operations periodically at an interval of \textit{64~ms} for each row, an interval that has been constant for the latest DRAM specifications [source] and called \textit{maximum refresh period}. This interval is conservatively halved if the DRAM exceeds the specified working temperature of $85^{\circ}C$.

### 2014-10-23 paper/include/discussion.tex

Another proposed technique is ESKIMO by Isen and John \cite{eskimo}, which focus at reducing total DRAM power through optimizing the accesses to the DRAM. By using one bit of a cache row to denote whether it hold nonsense data, it becomes possible to resolve some memory accesses in the caches, which otherwise would have needed to access DRAM. A memory region is regarded as nonsense if it has been deallocated, or allocated but not written to. The ISA has to be extended to provide the allocation and deallocation information. If a block has to be replaced in a write-through cache and the replaced block is dirty but also deallocated, the cache can ignore writing the replaced block to memory. Similarly, a write miss to a newly allocated area results in a read access that can also be ignored. Combining ESKIMO with the selective refresh implementation proposed by Ohsawa \cite{ohsawa} (ESKIMO[31]), $39\%$ of total DRAM power was saved on average.

\textbf{Broadly compare the techniques from different approaches}

>>>>

Another proposed technique is ESKIMO by Isen and John \cite{eskimo}, which focus at reducing total DRAM power through optimizing the accesses to the DRAM. By using one bit of a cache row to denote whether it hold nonsense data, it becomes possible to resolve some memory accesses in the caches, which otherwise would have needed to access DRAM. A memory region is regarded as nonsense if it has been deallocated, or allocated but not written to. The ISA has to be extended to provide the allocation and deallocation information. If a block has to be replaced in a write-through cache and the replaced block is dirty but also deallocated, the cache can ignore writing the replaced block to memory. Similarly, a write miss to a newly allocated area results in a read access that can also be ignored. Combining ESKIMO with the selective refresh implementation proposed by Ohsawa \cite{ohsawa}, $39\%$ of total DRAM power was saved on average.

\textbf{Broadly compare the techniques}
\begin{table*}[t]
	\caption{\label{tbl:summary}Summarized stuff.}
        \begin{center}
			\normalsize
			\begin{tabular}{ l r r r r r r}
				\textbf{Technique} & \parbox{2cm}{ \textbf{Refresh \linebreak information*}} & \textbf{Modifies} & \parbox{1.7cm}{ \textbf{Refresh \linebreak reduction}} & \parbox{2.2cm}{ \textbf{DRAM power \linebreak reduction}} & \parbox{1.2cm}{\textbf{Storage \newline Overhead}} & \parbox{1.5cm}{\textbf{Performance impact}} \vspace{0.05cm} \\
				\hline
				\textit{Smart Refresh} & A & MC & 59\% & $12.13\%$ & $0.0048\%$  & Unknown  \\
				\textit{Refrint} & A & MC, (?) & N/A & N/A  & $0.005\%$  & N/A  \\
				\hline
				\textit{RAIDR} & R & MC & $74.6\%$ & $8.3\%$  & $0.031\%$  & $4.1\%$  \\
				\textit{DTail} & V, R & MC, DDRx & $88.2\% \to 98.9\% $ & $< 23\%$  & $0.006\%$ & R:  \\
				\textit{RIO} & R & OS & $87.5\%$ & Unknown  & $0.1\%$  & $4.5\%$ \\
				\textit{SECRET} & R & MC & $87.5\%$ & $18.6\%$  & $>0.01\%$  & $\pm 1.4\%$  \\
				\hline
				\textit{PARIS} & V & MC & Unknown & $-41.7\%$ DTail-V  & $0.0015\%$  & $-4.4\%$ DTail-V  \\
				\hline
				\textit{Flikker} & T & MC, OS, Apps & Unknown & $20\% \to 25\%$  & $0.005\%$  & $-1\%$   \\
				\textit{Sparkk} & V, T  & MC, OS, Apps &  $+50\%$ Flikker & Unknown & Unknown  & Unknown \\
				\hline
			\end{tabular}
		\end{center}
\end{table*}

Smart refresh
Result shows that the refresh reduction was in average 59\%
and refresh power consumption savings was on average 52\%,
all without any performance degradation. 12.13\% reduction of total DRAM
The memory
footprint of the modifications is 0.0048\% of total device
capacity when three-bits counters are used assuming a DRAM
row size of 8 KB.

Refrint
N/A

Flikker
The partitioning scheme can potentially lower performance as
locality of data elements changes, but the found performance
degradation was 1\% or less. No loss of reliability was ex-
perienced, and the technique saved between 20-25\% of total
DRAM power consumption depending on application.
In simulations, N was two bits and Count
was 5 bits, giving a total overhead of 0.005\%, with a row size
of 8 KB.

Sparkk
Sparkk produces higher peak SNR at half the
average refresh rate used for Flikker.

Modifies applications:
Flikker, Sparkk

Modifies OS/runtime system:
ESKIMO, Flikker, Sparkk.

Modifies MC:
All but RIO and PARIS.

Only in SW:
RIO, PARIS.

ROR doesn't work with $<$ 32 GB, cant keep up the req. speed.

### 2014-10-23 paper/include/secret.tex

With SECRET Lin et al. \cite{secret} masks cells with short retention time using Error Correction Pointers (ECP). The DRAM cells are mapped through profiling, which is performed the first time the system is in use, and all cells that do not meet the retention time requirements is assigned an ECP. An ECP is basically a pointer to a cell that met the retention time requirements. The framework logic, which is implemented inside the MC, checks whether there is any ECP for the address accessed. If there is, it replaces the cell's data with the data from the cell pointed to by the ECP. The ECPs are stored in DRAM during runtime and are also cached in the MC to mask the delay of fetching them from DRAM.

>>>>

With SECRET Lin et al. \cite{secret} masks cells with short retention time (bad cells) using Error Correction Pointers (ECP). The DRAM cells are mapped through profiling, which is performed the first time the system is in use, and all bad cells are assigned an ECP. An ECP is basically a pointer to a cell that met the retention time requirements. The framework logic, which is implemented inside the MC, checks whether there is any ECP for the address accessed. If there is, it replaces the cell's data with the data from the cell pointed to by the ECP. The ECPs are stored in DRAM during runtime and are also cached in the MC to mask the delay of fetching them from DRAM.
